---
title: "R Case Study 1 - Retail"
output: pdf_document
geometry: margin=2cm
---


```{r include=FALSE}

#Packages
require(lubridate)
require(Hmisc)
require(dplyr)
require(reshape2)

# Import files 
Customer <- read.csv("Customer.csv", header = TRUE)
prod_cat_info <- read.csv("prod_cat_info.csv",header = T,na.strings = NA)
Transactions <- read.csv("Transactions.csv", header = T,)

#Type casting
Customer$customer_Id <- as.character(Customer$customer_Id)
Customer$city_code <- as.character(Customer$city_code)
Customer$DOB <- as.Date(Customer$DOB, format= "%d-%m-%Y")
Transactions$tran_date <- lubridate::dmy(Transactions$tran_date)
prod_cat_info$prod_cat_code <-as.character(prod_cat_info$prod_cat_code)   
prod_cat_info$prod_sub_cat_code <-  as.character(prod_cat_info$prod_sub_cat_code)
Transactions$transaction_id <- as.character(Transactions$transaction_id)
Transactions$cust_id <- as.character(Transactions$cust_id)
Transactions$prod_subcat_code <- as.character(Transactions$prod_subcat_code)
Transactions$prod_cat_code <- as.character(Transactions$prod_cat_code)


```


### Question 1. Merge the  Customer, prod_cat_info and Transactions Files as Customer_Final

a. Using base merge()
```{r}

# Q1 Merge the 3 Files as customer_final(Base merge)

Customer_Final <- merge(Transactions,Customer, by.x = "cust_id", 
                        by.y = "customer_Id",all.x = T)

Customer_Final <- merge(Customer_Final,prod_cat_info, by.x = "prod_cat_code", 
                        by.y = "prod_cat_code", all.x = T)


```


b. Using dplyr
```{r}

# Q1 Merge the 3 Files as customer_final (Dplyr merge)

Customer_Final <- dplyr::left_join(Transactions,Customer, by = c("cust_id" = "customer_Id"))

Customer_Final <- dplyr::left_join(Customer_Final,prod_cat_info, 
                                   by = "prod_cat_code")



```


### Question 2 Summary report
  a.Variable names and their data types of Customer_Final dataset

```{r}

#Q2 a. column names and corresponding data types
str(Customer_Final)

```

  b. Top 10 records of Customer_Final dataset

```{r}

#Q2 b.Top 10 observations
head(Customer_Final,10)

```


  c. 5-number summary of continuous variables

```{r}

#Q2 c.5 number summary for continuous variables

numericvariables <- Customer_Final[ ,sapply(Customer_Final, is.numeric)]
summary(numericvariables)

```


  d. Frequency tables for Categorical variables

```{r}

# Q2 d.Frequency table for all categorical variables

categoricalvariables <- Customer_Final[ ,sapply(Customer_Final, is.character)]
categoricalvars_F <- categoricalvariables[,-c(1,2)] #removing id variables
#Frequency table for categorical variables


Freq1<- data.frame(dplyr::group_by(categoricalvars_F,prod_subcat_code)%>%
                     summarise(Count = n())%>%arrange(prod_subcat_code))
Freq1

Freq2<- data.frame(dplyr::group_by(categoricalvars_F,prod_cat_code)%>%
                     summarise(Count = n()))%>%arrange(prod_cat_code)
Freq2

Freq3<-data.frame(dplyr::group_by(categoricalvars_F,Store_type)%>%
                    summarise(Count = n()))%>%arrange(Store_type)
Freq3

Freq4<- data.frame(dplyr::filter(categoricalvars_F,Gender != "")
                   %>%group_by(Gender)
                   %>%summarise(Count = n()))%>%arrange(Gender)
Freq4

Freq5<- data.frame(dplyr::filter(categoricalvars_F,city_code != "" )%>%
                     group_by(city_code)%>%
                     summarise(Count = n()))%>%arrange(city_code)
Freq5

Freq6 <-data.frame(dplyr::group_by(categoricalvars_F,prod_cat)%>%
                     summarise(Count = n()))%>%arrange(prod_cat)
Freq6

Freq7<- data.frame(dplyr::group_by(categoricalvars_F,prod_sub_cat_code)
                         %>%summarise(Count = n())
                   %>%arrange(prod_sub_cat_code))
Freq7

Freq8<- data.frame(dplyr::group_by(categoricalvars_F,prod_subcat)%>%
                       summarise(Count = n()))%>%arrange(prod_subcat)
Freq8


```


### Question 3. Graphical representation of variables
  a. Histogram for numeric variables
```{r}

#Q3 1.Histogram for all the numeric variables

hist1 <- hist(numericvariables$Qty,xlab = "Qty",main="Histogram for Qty")

hist2 <- hist(numericvariables$Rate,xlab = "Rate",main="Histogram for Rate")

hist3 <- hist(numericvariables$Tax,xlab = "Tax",main="Histogram for Tax")

hist2 <- hist(numericvariables$total_amt,xlab = "total_amt",
              main="Histogram for total_amt")

```


  b. Bar Graph for categorical variables
```{r}

#Bar Graph for all the Categorical variables

freqbar1<- barplot(Count~prod_subcat_code,Freq1,xlab = "prod_subcat_code",
                   col.lab = "Red",main = "Bar Chart - prod_subcat_code ")

freqbar2 <- barplot(Count~prod_cat_code,Freq2,xlab = "prod_cat_code",
                    col.lab = "Red",main = "Bar Chart - prod_cat_code ")

freqbar3 <- barplot(Count~Store_type,Freq3,xlab = "Store_type",col.lab = "Red",
                    main = "Bar Chart - Store_type ")

freqbar4 <- barplot(Count~Gender,Freq4,xlab = "Gender",col.lab = "Red",
                    main = "Bar Chart - Gender ")

freqbar5 <- barplot(Count~city_code,Freq5,xlab = "city_code",col.lab = "Red",
                    main = "Bar Chart - city_code ")

freqbar6 <- barplot(Count~prod_cat,Freq6,xlab = "prod_cat",col.lab = "Red",
                    main = "Bar Chart - prod_cat ")

freqbar7 <- barplot(Count~prod_sub_cat_code,Freq7,xlab = "prod_sub_cat_code",las=2,
                     col.lab = "Red",main = "Bar Chart - prod_sub_cat_code ")

freqbar8 <- barplot(Count~prod_subcat,Freq8,xlab = "prod_subcat",las=2,
                     col.lab = "Red",main = "Bar Chart - prod_subcat ")


```


### Question 4. Information from data

  a.Time period for the transaction data

```{r}

#Q4 a. Calculate the following time period of the available transaction data

firstdate <- min(Customer_Final$tran_date)
lastdate <- max(Customer_Final$tran_date)
paste(as.numeric(lastdate - firstdate),"days")

```


  b. Number of transactions where the total amount of the transactions are negative

```{r}

#Q4 b. Count the transactions where the total amount of the transactions 
#are negative

dplyr::filter(Customer_Final,total_amt< 0)%>%summarise(Count = n())


```


### Question 5. Product categorie/s that are more popular amongst F customers than male customers 

```{r}

#Q5 Analyze which product categories are more popular among female vs male 
#customers
Categories <- data.frame(dplyr::filter(Customer_Final,Gender != "")%>%
                           group_by(prod_cat,Gender)%>%
                           summarise(frequency = n()))
ProdGender <- reshape2::dcast(Categories,prod_cat~Gender)
ProdGender[ProdGender$F>ProdGender$M,1]

```


### Question 6. City Code that has the maximum customers and the percentage of customers from that city.

```{r}

#Q6 Which city code has the maximum customers and what was the percentage of 
#customers from that city.
sub1<- data.frame(dplyr::filter(Customer_Final,city_code != "")%>%
                    group_by(city_code,cust_id)%>%
                    summarise(frequency = n()))
sub1<-data.frame(group_by(sub1,city_code)%>%
                   summarise(noofcustomers = n()))
sub1$percentage <- round((sub1$noofcustomers/sum(sub1$noofcustomers))*100,2)
sub1[sub1$noofcustomers == max(sub1$noofcustomers),]

```


### Question 7. Store type that sells maximum products by value and by quantity

```{r}

#Q7Which store type sells maximum product by value and by quantity
#By Quantity
MAxprodbyQTY <- dplyr::group_by(Customer_Final,Store_type)%>%
  summarise(TProdSold = sum(Qty))
MAxprodbyQTY[MAxprodbyQTY$TProdSold == max(MAxprodbyQTY$TProdSold),1]

#By Value
MAxprodbyVALUE <- dplyr::group_by(Customer_Final,Store_type)%>%
  summarise(TProdSold = sum(total_amt))
MAxprodbyVALUE[MAxprodbyVALUE$TProdSold == max(MAxprodbyVALUE$TProdSold),1]


```


### Question 8. Total amount earned for Electronics and Clothing categories from Flagship stores.

```{r}

#Q8What was the total amount earned from electronics and clothing categories 
#from Flagship stores
ProdTamt <- data.frame(dplyr::filter(Customer_Final,Store_type == "Flagship store")
                       %>% group_by(prod_cat)
                       %>% summarise(Totalamtearned=sum(total_amt)))
ProdTamt[ProdTamt$prod_cat=="Electronics"|ProdTamt$prod_cat=="Clothing",]

```


### Question 9. Total amount earned from Male customers under the Electronics category

```{r}

#Q9 What was the total amount earned from male customers under the electronics
#category
data.frame(dplyr::filter(Customer_Final,prod_cat=="Electronics")%>%
             group_by(Gender)%>%
             summarise(Totalamtearned=sum(total_amt)))%>%
  filter(Gender == "M")

```


### Question 10.Number of customers who have more than 10 unique transactions after removing all transactions which have any negative amounts

```{r}

#Q10 How many customers have more than 10 unique transactions after removing all transactions which have any negative amounts 
A <- data.frame(dplyr::filter(Customer_Final,total_amt >0)%>%
                  group_by(cust_id,transaction_id)%>%
                  summarise(count = n()))
B <- data.frame(table(A$cust_id))
length( B[B$Freq>10,1])

```


### Question 11. For all customers ages 25-35
  a. Total amount spend for "Electronics and "Books" product categories
  
```{r}

#Q11 a.For all customers ages 25-35 calculate what was the total amount spent
#for "Electronics and "Books" product categories
Customer_Final$Age <- round(as.numeric(Sys.Date() - Customer_Final$DOB )/365.25)

ProdTamt2 <- data.frame(dplyr::filter(Customer_Final, Age >= 25 & Age <= 35)
                        %>%group_by(prod_cat)%>% 
                          summarise(TotalamtSpent=sum(total_amt)))
ProdTamt2[ProdTamt2$prod_cat=="Electronics"|ProdTamt$prod_cat=="Books",]

```


  b. Total amount spent by these customers between 1st Jan,2014 to 1st Mar,2014
  
```{r}

#Q11 b.For all customers ages 25-35 calculate what was the total amount spent by
#these customers between 1st Jan,2014 to 1st Mar,2014
data.frame(dplyr::filter(Customer_Final, Age >= 25 & Age <= 35,
                         tran_date >= "2014-01-01"& tran_date <= "2014-03-01")
           %>%summarise(TotalamtSpent=sum(total_amt)))

```

